---
title: "Aids Drug"
author: "santiago hernandez"
date: "11/17/19"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
avast <-read.csv(file="avastivironal.csv",header=TRUE)
avastf <- filter(avast,sex=="gal")
avastg <- filter(avast,LGBTQ=="gay")
avastm <- filter(avast,sex=="guy")
avasts <- filter(avast,LGBTQ=="straight")
avastd <- filter(avast,DrugUser=="yes")

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides (abacavirinol) is under test in clinical trials. Analyze the effectiveness of this drug across the board. Consider population subgroups by gender, drug use, and sexual orientation. Finish this! 

Instructions: We are going to have 5 different pie square test, one for each variabe.
In the readme section on github you need to explain that you are going to do five different reports, to make five different cases in each of the reports. 
 
Question: Determine how the drug should be working with respect to each subgroup. 

# 1, everyone

 How do we determine how the drug should be working with respect to the infected and the treatment group?
 
 \(H_0\):There is not a difference in the indivduals who are infected to AIDS and to those independent of their subgroup.

\(H_A\):There is a statistical difference between the indivuals who a are infected to AIDS and to those independent to their subgroup


# Methods

i will be using both descriptive and numerical and graphical results.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=avast,type="percent")
```
 
The data shows of Everyone taking the AID preventive drug 40% contracted AIDS and those taking the placebo 60% contracted AIDS. 

On the left hand side of the graph it shows that data for Everyone who did not contract AIDS while taking the Aids preventive drug and that was 50% and those taking the placebo stated 50 - 52%

```AIDS preventive drug
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment , data=avast)
rowPerc(table1)
colPerc(table1)
```
the drug is in full effect because the people who have consumed the drug only 41% contracted AIDS, while those taking the placebo 59% contracted AIDS. thus by common sense , those who took the AIDS drug have a less chance than those with a placebo.

the data shows 50/50 for those taking the drug and placebo. There wasn't a huge difference. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```
the drug is useful and effective because a large portion did not contract aids

portion infected:504.15+530.85=1035

portion not infected:10086.85+10621.15=30,794

Total effectiveness:10086.85+504.15=10086.85/10590.85=95%

those who took the drug showed 95% effectivness.


# 2,women

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

In question 2 our goal is to determine how the AID preventive drug should be working with respect to the sex of the women. 

 \(H_0\):There is no difference in the AID preventive drug that is independent to the sex of women.

\(H_A\):There is a statistical difference between AID preventive drug  and to those independent to the sex of the women. 




## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

i will be using both graphical descriptive and numerical descriptive results.

### Descriptive Results


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

i use the bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~infected + treatment,data=avastf,type="percent")
```

The data shows of the Women taking the AIDS drug were at 35% contracted aids and those who took the placebo had a toll of 65% contracted AIDS. 

the left side shows that data for the Men who did not get AIDS while taking the Aids preventive drug and that was 50% and those taking the placebo stated 48%.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=avastf)
rowPerc(table1)
colPerc(table1)
```

The drug is effective because for those taking the Aid preventive drug only 31% contracted AIDS, while those taking the placebo 69 percent contracted AIDS. taking the real AIDS preventive drugs has individuals at lower rates of contracting aids.
For those indivduals that didnt contract AIDS, the data is 50/50 for those taking the drug and placebo.  

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The drug is effective due to a large majority of indviduals that did not contract AIDS.

portion infected:332.85+345.15=678

portion not infected:4564.15+4732.85=9297

Toll of effectiveness: 4564.15+332.85=4564.15/4897=93%

the total who took the drug results showed 93% effectivness.

#   3,men

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

In question 3 our goal is to determine how the AID preventive drug should be working with respect to the sex of the males. 

 \(H_0\):There is no difference in the AID preventive drug that is independent to the sex of the men.

\(H_A\):There is a statistical difference between AID preventive drug  and to those independent to the sex of the men. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

i will be using both graphical descriptive and numerical descriptive results.

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

i use bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~infected+ treatment,data=avastm,type="percent")
```

The data shows of the Men taking the AID preventive drug 60% contracted AIDS and those taking the placebo 40% contracted AIDS. 

the left side of the grapgh shows those who took the AIDS preventive drug were at 50 percent and those whom consumed the placebo are at 45%
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=avastm)
rowPerc(table1)
colPerc(table1)
```

the drug is effective because those who took the AIDS preventive drug contracted only 59 percent contracted aids, and those who took the placebo contracted 41 percent.
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The drug is effective due to a large majority of indviduals that did not contract AIDS.

poriton of Those infected:172.74+184.26=357

portion of Those not infected:5521.26+5889.74=11,411

Toll of effectiveness: 5521.26+172.74=5521.26/5694=97%

the total who took the drug results showed 97% effectivness.

#4, gay individuals

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

In question 3 our goal is to determine how the AID preventive drug should be working with respect to gay indivduals. 

\(H_0\):There is no difference in the AID preventive drug that is independent to the LGBTQ of the gay indivduals.

\(H_A\):There is a statistical difference between AID preventive drug  and to those independent to the LGBTQ of the gay indivduals. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

i will be using both graphical descriptive and numerical descriptive results.

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~infected + treatment,data=avastg,type="percent")
```
out of the gay indviduals the data suggest of the people who took the drug 70% contracted aids while only 30% who took the placebo contracted AIDS. 

the right side of the  graph it shows that of the gay indviduals that to the AIDS preventive drug 50% did not contract AIDS, and those taking the placebo 55% did not get AIDS. 
We can conclude the drug is effective.
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=avastg)
rowPerc(table1)
colPerc(table1)

```

72% of those who took the AIDS preventive drug contracted AIDS and while those who took the placebo only 28% contracted AIDS.
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The drug is effective due to a large majority of indviduals not contracting aids.

portion of Those infected: 52.69+52.31=105

portion of Those not infected:1020.31+1022.69=2043

Toll of effectiveness:1030.31+52.69=1030.31/1083=95%

the total who took the drug results showed 95% effectivness.

#  5,straight invividuals

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

In question 5 our goal is to determine how the AID preventive druh should be working with respect to to each straight indvidual.

\(H_0\):There is no difference in the AID preventive drug that is independent to the LGBTQ of the straight indivduals.

\(H_A\):There is a statistical difference between AID preventive drug  and to those independent to the LGBTQ of the straight indivduals. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

i will be using both graphical descriptive and numerical descriptive results.

### Descriptive Results



#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We use bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~infected + treatment,data=avasts,type="percent")
```

Of the straight indviduals, the data suggest the people who took the AIDS preventive drug  38% contracted aids while 62% who took the placebo also contracted AIDS. 

On the right side of the  graph it states of the straight indviduals that to the AIDS preventive drug 50% did not contract AIDS, and those taking the placebo 55% did not contract AIDS. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=avasts)
rowPerc(table1)
colPerc(table1)
```
It states 37% of those who took the AIDS preventive drug contracted AIDS and while those who took the placebo 63% contracted AIDS.

the indivduals that didnt contract AIDS, the data is 50/50 for those taking the drug and placebo.  
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The drug is effective due to a large majority of indviduals that did not contract AIDS.

portion of Those infected: 451.49+478.51=930

portion of Those not infected:9056.51+9598.49=18,655

Toll of effectivenss: 9056.51+451.49=9056/9508=95%

Out of the total who took the drug results showed 95% effectivness.

# 6, the drug users

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

In question 6,  our goal is to determine how the AID preventive drug should be working with respect to each DrugUser.

\(H_0\):There is no difference in the AID preventive drug that is independent to DrugUsers.

\(H_A\):There is a statistical difference between AID preventive drug  and to those independent to DrugUsers.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

i will be using both graphical descriptive and numerical descriptive results.

### Descriptive Results


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

i use bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~infected + treatment,data=avastd,type="percent")
```

The data shows of the DrugUsers taking the AID preventive drug 25% contracted AIDS and those taking the placebo 65% contracted AIDS. 

on the left of the grapgh it shows that only 25 percent contracted aids and 66 who took the placebo contracted aids.
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=avastd)
rowPerc(table1)
colPerc(table1)
```

The drug is taking effect because for those taking the Aid preventive drug only 24 percent contracted aids while those who took the placebo only 76 percent contracted aids thus data shows the aids preventive drugs is taking effect.
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The drug is effective due to a large majority of indviduals that did not contract AIDS.

portion of people infected:64.82+75.18=140

portion of people not infected:1112.18+1289.82=2401

toll of effectiveness: 1112.18+64.82=1112.18/1177=95%

Out of the total who took the drug results showed 95% effectivness.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In conclusion to all 6 of the subgroups there is no correlation based on all the subgroups combined that causes one subgroup or the other that increases there odds of contracting AIDS.

the data presented mostly shows that the AIDS preventive drug was effective.

Based on the chi-square data on all of the subgroups the AIDS preventive drug proved effective ranging from 93-96% effective. 

In conclusion the real pill has a lower rate of contracting AIDS

